title: "GAP workflow with packaged example data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GAP workflow with packaged example data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
`GAP` fits a genetic association with progression model using summary-level
effect sizes and standard errors.
This vignette shows how to:
1. Load the packaged example input file.
2. Inspect the required columns.
3. Run a small prior-estimation example.
4. Use the fitted prior in `GAP_bayesian_lrt()`.
## Load the example data
```{r}
library(GAP)
gap_input <- gap_example_data()
dim(gap_input)
names(gap_input)
head(gap_input)
```
The bundled example file path is also available through:
```{r}
gap_example_path()
```
## Estimate prior parameters
```{r}
input <- gap_input
prior_params <- GAP_bayesian_prior(
  input = input,
  alpha = log10(0.05),
  p_threshold = 5e-8,
  random = 50
)
str(prior_params)
```
After estimating the prior, you can pass it to `GAP_bayesian_lrt()` on the
dataset you want to analyze.
```{r, eval = FALSE}
lrt_results <- GAP_bayesian_lrt(
  input = input,
  alpha = log10(0.05),
  prior = prior_params
)
head(lrt_results)
```
