title: "GAP workflow with packaged example data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GAP workflow with packaged example data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`GAP` fits a genetic association with progression model using summary-level
effect sizes and standard errors.

This vignette shows how to:

1. Load the packaged example input file.
2. Inspect the required columns.
3. Run a small prior-estimation example.
4. Use the fitted prior in `GAP_bayesian_lrt()`.

## Load the example data

```{r}
library(GAP)

gap_input <- gap_example_data()

dim(gap_input)
names(gap_input)
head(gap_input)
```

The bundled example file path is also available through:

```{r}
gap_example_path()
```

## Estimate a prior on a small subset

`GAP_bayesian_prior()` can be expensive on large inputs, so the vignette uses a
small subset just to show the workflow.

```{r}
set.seed(1)
subset_idx <- sample(seq_len(nrow(gap_input)), 250)
prior_example <- GAP_bayesian_prior(
  input = gap_input[subset_idx, ],
  alpha = 0,
  p_threshold = 1e-4,
  random = 100
)

str(prior_example)
```

After estimating the prior, you can pass it to `GAP_bayesian_lrt()` on the
dataset you want to analyze.

```{r, eval = FALSE}
lrt_results <- GAP_bayesian_lrt(
  input = gap_input,
  alpha = 0,
  prior = prior_example
)

head(lrt_results)
```
