---
title: "GAP workflow with packaged chr22 example data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GAP workflow with packaged chr22 example data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`GAP` fits a genetic association with pleiotropy model using summary-level
effect sizes and standard errors.

This vignette shows how to:

1. Load the packaged example files.
2. Parse SNP, chromosome, and position fields from the chr22 example IDs.
3. Build a `GAP_bayesian_prior()` input table.
4. Run a small prior-estimation example.

## Load the example files

```{r}
library(GAP)

example_files <- gap_example_data()
beta <- example_files$beta
se <- example_files$se

dim(beta)
names(beta)[1:5]
head(beta[, 1:4])
```

The bundled files are the original gzipped tables under `inst/extdata`, so the
same paths are also available through:

```{r}
gap_example_beta_path()
gap_example_se_path()
```

## Build a GAP input table

The example IDs include chromosome and genomic position in the form
`gene-chr:pos_ref/alt`. The helper code below extracts the pieces needed by the
GAP functions.

For illustration, we use:

- `Astrocytes` as `beta_01`
- `Excitatory.neurons` as `beta_12`
- `bulk` as `beta_cc`

You can swap these columns for the pair of cell types relevant to your
analysis.

```{r}
parse_chr22_id <- function(x) {
  chr_pos <- sub("^.*-([0-9]+):([0-9]+)_.*$", "\\1\t\\2", x)
  parts <- strsplit(chr_pos, "\t", fixed = TRUE)
  data.frame(
    chr = as.integer(vapply(parts, `[`, "", 1)),
    pos = as.integer(vapply(parts, `[`, "", 2))
  )
}

loc <- parse_chr22_id(beta$ID)

gap_input <- data.frame(
  snp = beta$ID,
  chr = loc$chr,
  pos = loc$pos,
  beta_01 = beta$Astrocytes,
  beta_12 = beta$Excitatory.neurons,
  beta_cc = beta$bulk,
  se_01 = se$Astrocytes,
  se_12 = se$Excitatory.neurons,
  se_cc = se$bulk,
  check.names = FALSE
)

head(gap_input)
```

## Estimate a prior on a small subset

`GAP_bayesian_prior()` can be expensive on large inputs, so the vignette uses a
small subset just to show the workflow.

```{r}
set.seed(1)
subset_idx <- sample(seq_len(nrow(gap_input)), 250)
prior_example <- GAP_bayesian_prior(
  input = gap_input[subset_idx, ],
  alpha = 0,
  p_threshold = 1e-4,
  random = 100
)

str(prior_example)
```

After estimating the prior, you can pass it to `GAP_bayesian_lrt()` on the
dataset you want to analyze.

```{r, eval = FALSE}
lrt_results <- GAP_bayesian_lrt(
  input = gap_input,
  alpha = 0,
  prior = prior_example
)

head(lrt_results)
```
